New lower bounds for Schur and weak Schur numbers

### 2021-04-16 article/article.tex
amelioration introduction

\usepackage[table]{xcolor}
\usepackage[a4paper, total={6in, 8in}]{geometry}

@@

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by continuing on
Rowley's "template"-based approach for Schur and Ramsey numbers. Finding suitable templates allows us to apply Rowley's construction
to get explicit partitions improving lower bounds as well as the growth rate for both Schur numbers and Ramsey numbers \(R_n(3)\).
We also developed a new method to improve lower bounds on weak Schur numbers. Furthermore, this paper tries to analyze former
works on the subject based on the principle that good partitions into \(n+1\) subsets start with good partitions
into \(n\) subsets. We show that exceeding the previous lower bound \(\WS (6) \geqslant 582\) is impossible with such an assumption upon
imposing certain conditions on the good 5-subsets partition. The new lower bounds include \(S(9) \geqslant 17803\), \(S(10) \geqslant 60948\),
\(\WS (9) \geqslant 22536 \) and \(\WS (10) \geqslant 71214 \).

>>>>

\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage[a4paper, total={6in, 8in}]{geometry}

@@

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. Finding templates allows us to 
apply Rowley's construction and to get explicit partitions improving lower bounds as well as the growth rate for 
both Schur numbers and Ramsey numbers \(R_n(3)\). We also developed a new method to improve lower bounds on 
weak Schur numbers. Furthermore, this paper tries to analyze former works on the subject based on the principle 
that good partitions into \(n+1\) subsets start with good partitions into \(n\) subsets. We show that exceeding the 
previous lower bound \(\WS (6) \geqslant 582\) is impossible with such an assumption upon imposing certain conditions 
on the good 5-subsets partition. The new lower bounds include \(S(9) \geqslant 17803\), \(S(10) \geqslant 60948\),
\(\WS (9) \geqslant 22536 \) and \(\WS (10) \geqslant 71214 \).

### 2021-04-16 article/introduction.tex
amelioration introduction

We are interested in partioning the set of integers \(\{1, ..., p\}\) in \(n\) subsets such that there is no subset 
containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are sum-free. If we 
add the hypothesis \(x \neq y\), we say the subsets are weakly sum-free. The greatest \(p\) for which a 
partition into \(n\) sum-free exists is called the \(n^{\text{th}}\) Schur number and is denoted \(S(n)\) 
\cite{Schur1917}. Likewise for weakly sum-free partitions we define \(WS(n)\) the \(n^{\text{th}}\) weak Schur 
number \cite{Irving1973}. Only the first values of these sequences are known. This article concerns the lower 
bounds of these numbers.

\par Before Rowley's "template"-based approach for Schur and Ramsey numbers \cite{RowleyRamsey}, the 
previous generic construction for Schur numbers was given by Abbott and Hanson \cite{AbbottHanson} in 1972 
with a recursive construction. It gave the best lower bounds for all sufficiently large bounds. No equivalent 
was known for weak Schur numbers and as a result the best known partitions for large weak Schur numbers 
did not utilize the weakly sum-free hypothesis. As for smaller numbers, the best lower bounds were obtained 
by conducting a computer search. This search for weakly sum-free partitions relied on the recursive assumption 
that a good weakly sum-free partition into \(n+1\) colors starts with a good weakly sum-free partition into \(n\) 
colors. In 2020, Rowley introduced the notion of template for Schur and Ramsey numbers which generalizes 
Abbott and Hanson's construction and gives new lower bounds (and inequalties) for Schur numbers. Rowley also 
gives two inequalities for weak Schur numbers that yield significant improvements over previous lower bounds 
which besides do utilize the weakly sum-free hypothesis \cite{RowleyWS}.

\renewcommand{\arraystretch}{1.5}

@@

\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    old & 1 & 4 & 13 & 44 & 160 & 536 & 1680 & 5041 & 15124 & 51120 & 172216 & 575664 \\
      & & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{ELIAHOU2012175} & 
    \cite{ELIAHOU2012175} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    Rowley & & & & & & & & 5286 & 17694 & 60320 & 201696 & 631840 \\
    \hline
    new & & & & & & & & & 17803 & 60948 & 203828 & 638548 \\
    \hline
\end{tabular}

\vspace{3ex}
\textbf{Table 2 - Comparison of lower bounds for weak Schur numbers}

\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    old & 2 & 8 & 23 & 66 & 196 & 582 & 1740 & 5201 & 15596 & 51520 & 172216 & 575664 \\
      & & & & & \cite{ELIAHOU2012175} &\cite{EliahouBook} & \cite{Rafilipojaona} & \cite{Rafilipojaona} & 
    \cite{Rafilipojaona} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    Rowley & & & & & & 642 & 2146 & 6976 & 21848 & 70778 & 241282 & 806786 \\
    \hline
    new & & & & & & & & & 22536 & 71214 & 243794 & 815314 \\
    \hline
\end{tabular}

@@

\par The main contribution of this article is a generalization of the concept of template to weak Schur numbers. 
This gives new lower bounds (and inequalities) for weak Schur numbers. This construction also includes as a 
special case an analogous for weak Schur numbers of Abbott and Hanson's construction.

\par We first explain Rowley's template-based construction in the context of Schur numbers and then give 
new templates, thus providing new lowers bounds and inequalities as well as showing that the growth rates 
for both Schur and Ramsey numbers exceed 3.28. Then, we generalize the concept of templates to weak Schur 
numbers and provide new lower bounds for weak Schur numbers. Finally, we analyze the significant difference 
between new lower bounds obtain with templates and the former lower bounds obtained by computer search 
and we provide evidence which indicate that the main assumption made in those articles removes the optimal 
partitions from the search space.

@@

\begin{definition}
We sometimes refer to a partitition as a colouring. The colouring associated to a partition \(A_1, ..., A_n\) of
\([\![1, p]\!]\) is the function \(f\) such that \(\forall x \in [\![1, p]\!], x \in A_{f(x)}\). Likewise, the partition 
associated to a colouring \(f\) of \([\![1, p]\!]\) with \(n\) colors is \(\forall c \in [\![1, n]\!], A_c = f^{-1}(c)\).
\end{definition}

>>>>

\qquad We are interested in partioning the set of integers \(\{1, ..., p\}\) in \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Only the first values of these sequences are known. 
This article focuses on the lower bounds for these numbers.

\par
Before Rowley's "template"-based approach for Schur and Ramsey numbers \cite{RowleyRamsey}, the 
previous generic construction for Schur numbers was given by Abbott and Hanson \cite{AbbottHanson} in 1972 
with a recursive construction. It used to give the best lower bounds for all sufficiently large numbers. No equivalent 
was known for weak Schur numbers and as a result the best known partitions for large weak Schur numbers 
did not utilize the weakly sum-free hypothesis. As for smaller numbers, the best lower bounds were obtained 
by conducting a computer search. This search for weakly sum-free partitions relied on the recursive assumption 
that a good weakly sum-free partition into \(n+1\) colors starts with a good weakly sum-free partition into \(n\) 
colors. In 2020, Rowley introduced the notion of template for Schur and Ramsey numbers which generalizes 
Abbott and Hanson's construction and gives new lower bounds (and inequalties) for Schur numbers. Rowley also 
gives two inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements over previous 
lower bounds which besides do utilize the \textit{weakly} sum-free hypothesis.

\renewcommand{\arraystretch}{1.2}

@@

\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{old} & 1 & 4 & 13 & 44 & 160 & 536 & 1680 & 5041 & 15124 & 51120 & 172216 & 575664 \\
      & & & & & \cite{Heule2017} & \cite{Fredricksen} & \cite{Fredricksen} & \cite{ELIAHOU2012175} & 
    \cite{ELIAHOU2012175} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyRamsey}\end{tabular} & & & & & & & & 5286 & 17694 & 
    60320 & 201696 & 631840 \\
    \hline
    \hyperref[Schur]{new}  & & & & & & & & & 17803 & 60948 & 203828 & 638548 \\
    \hline
\end{tabular}

\vspace{3ex}
\setlength{\tabcolsep}{4pt}

\textbf{Table 2 - Comparison of lower bounds for weak Schur numbers}

\begin{tabular}{|*{13}{c|}}
    \hline
    \(n\) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 \\
    \hline
    \multirow{2}{*}{old} & 2 & 8 & 23 & 66 & 196 & 582 & 1740 & 5201 & 15596 & 51520 & 172216 & 575664 \\
      & & & & & \cite{ELIAHOU2012175} &\cite{EliahouBook} & \cite{Rafilipojaona} & \cite{Rafilipojaona} & 
    \cite{Rafilipojaona} & \cite{AbbottHanson} & \cite{AbbottHanson} & \cite{AbbottHanson} \\
    \hline
    \begin{tabular}{@{}c@{}}Rowley \\ \cite{RowleyWS}\end{tabular} & & & & & & 642 & 2146 & 6976 & 21848 
    & 70778 & 241282 & 806786 \\
    \hline
    \hyperref[WeakSchur]{new} & & & & & & & & & 22536 & 71214 & 243794 & 815314 \\
    \hline
    \(\displaystyle \frac{\text{new} - \text{old}}{\text{old}}\) & 0\% & 0\% & 0\% & 0\% & 0\% & 10.3\% & 
    23.3\% & 34.1\% 
    & 44.5\% & 38.2\% & 41.6\% & 41.6\% \\
    \hline
\end{tabular}

@@

\par
The main contribution of this article is a generalization of the concept of template to weak Schur numbers. 
This gives new lower bounds (and inequalities) for weak Schur numbers. This construction also includes as a 
special case an analogous for weak Schur numbers of Abbott and Hanson's construction for Schur numbers.

\par
We first explain Rowley's template-based construction in the context of Schur numbers and then give 
new templates, thus providing new lowers bounds and inequalities as well as showing that the growth rates 
for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
We then  generalize the concept of templates to weak Schur numbers and provide new lower bounds for weak 
Schur numbers.

\par
Finally, we analyze the significant difference between new lower bounds obtain with templates and the former 
lower bounds obtained by computer search and we provide evidence which indicate that the main assumption 
made in those articles removes the optimal partitions from the search space.

\par
We now introduce notations and definitions we use throughout this article.

@@

\begin{definition}
A partitition is sometimes referred to as a colouring. The colouring associated to a partition \(A_1, ..., A_n\) of
\([\![1, p]\!]\) is the function \(f\) such that \(\forall x \in [\![1, p]\!], x \in A_{f(x)}\). Likewise, the partition 
associated to a colouring \(f\) of \([\![1, p]\!]\) with \(n\) colors is \(\forall c \in [\![1, n]\!], A_c = f^{-1}(c)\).
\end{definition}
